---
title: "Untitled"
output: html_document
---

# Load processed datasets
```{r}
train2 <- readRDS("../data/train.movierank.RData")
test2 <- readRDS("../data/test.movierank.RData")
load("../data/MS_sample/train1.RData")
load("../data/MS_sample/test1.RData")
```

# Model-based Algorithm

### EM algorithm 
```{r}

N <- dim(train1)[1]
D <- dim(train1)[2]
N2 <- dim(test1)[1]

I_1 = list() # list of papers voted 1 by each user 
I_1[[1]] = as.numeric(unlist(which(train1[1,] == 1)))
for (i in 2:N) {
  k = as.numeric(unlist(which(train1[i,] == 1)))
  I_1 = c(I_1,list(k))
}

I_0 = list() # list of papers voted 0 by each user 
I_0[[1]] = as.numeric(unlist(which(train1[1,] == 0)))
for (i in 2:N) {
  k = as.numeric(unlist(which(train1[i,] == 0)))
  I_0 = c(I_0,list(k))
}

J_1 = list() # list of users who vote 1 each paper
J_1[[1]] = as.numeric(unlist(which(train1[,1] == 1)))
for (i in 2:D) {
  k = as.numeric(unlist(which(train1[,i] == 1)))
  J_1 = c(J_1,list(k))
}

J_0 = list() # list of users who vote 0 each paper
J_0[[1]] = as.numeric(unlist(which(train1[,1] == 0)))
for (i in 2:D) {
  k = as.numeric(unlist(which(train1[,i] == 0)))
  J_0 = c(J_0,list(k))
}

multinomialEM <- function(data, C, t = 1*10^-6){
  change_1 <- 1 # Measures change in centroids
  change_0 <- 1
  #cp <- matrix(1/C, nrow = N, ncol = C)
  cp <- runif(C)
  cp = cp/sum(cp)
  gamma_1 = matrix(runif(C*D), nrow = C, ncol = D)
  gamma_0 = 1-gamma_1
  A <- matrix (0, N, C) # responsibilities pi
  Phi <- matrix(0, N, C)

  while(change_1 > t | change_0 > t){
    # E step
    for (k in 1:C) {
      for (i in 1:N) {
        Phi[i,k] = prod(gamma_0[k,I_0[[i]]]) * prod(gamma_1[k,I_1[[i]]]) * cp[k]
      }
    }
    
    A<-Phi/rowSums(Phi)
    
    #for (i in 1:N) {
    #  A[i, ] <- Phi[i, ]/sum(Phi[i, ]) # responsibilities pi
    #}
    
    # M step
 
    cp = colSums(A)/N
    
    
    gamma_1_old = gamma_1
    gamma_0_old = gamma_0
    
    for (k in 1:C) {
      for (j in 1:D) {
        gamma_1[k,j] = sum(A[J_1[[j]],k])/sum(A[,k])
        gamma_0[k,j] = sum(A[J_0[[j]],k])/sum(A[,k])
      }
    }
    
    change_1 = norm(gamma_1_old - gamma_1, type = "O")
    change_0 = norm(gamma_0_old - gamma_0, type = "O")
  }
  return(result = list(cp, gamma_1, gamma_0, A))

}

```

### Predict after training
```{r}
# Prediction
predictCF <- function(data_test = test1, param){
  B = list() # list of papers needed to predict from each user
  B[[1]] = as.numeric(unlist(which(data_test[1,] == 1)))
  for (i in 2:N2) {
    k = as.numeric((which(data_test[i,] == 1)))
    B = c(B,list(k))
  }

  gamma1 = param[[2]]
  A = param[[4]]
  expect.mat = matrix(0, ncol = ncol(data_test), nrow = nrow(data_test))
  colnames(expect.mat) = colnames(data_test)
  for (i in 1:length(B)) {
    for (j in 1:length(B[[i]])) {
      if(length(B[[i]]) == 0){
        expect.mat[i,B[[i]][j]] = NA
    }
      else{
        expect.mat[i,B[[i]][j]] = (A[i,] %*% gamma1[,B[[i]]])[j]
    }

   }
 }

  return(expect.mat)
}

```

### Use CV to pick the best cluster number and use ranked scoring for evaluation
```{r}
# Separate the training data into two parts--training set and validation set
  
   training_data = train1
   validation_data = matrix(0,nrow = nrow(train1), ncol = ncol(train1))
   rownames(validation_data)<-rownames(training_data)
   colnames(validation_data)<-colnames(training_data)
   set.seed(2249)
   index<-c()
   
    for (i in 1:nrow(training_data)) {
      
      index<-as.numeric(unlist(which(train1[i,] == 1)))
      v_index <- sample(index, round(length(index) * 0.25))
      training_data[i,v_index]<-0
      validation_data[i,v_index]<-1
      
    }

# Caculate the accuracy rate using ranked scoring
   
R<-function(predict,true,d=0,alpha=10){
  
  r<-rep(0,nrow(true))
  r_max<-rep(0,nrow(true))
  
  for(a in 1:nrow(true)){
    
    r[a]<-sort(r[a],decreasing = T)
    true[a,]<-sort(true[a,],decreasing = T)
    
    for(j in 1:ncol(true)){
      
      r[a]<-r[a]+max(predict[a,]-d,0)/2^((j-1)/(alpha-1))
      r_max[a]<-r_max[a]+sum(max(true[a,]-d,0)/2^((j-1)/(alpha-1)))
      
    }
  }
  
  rs<-100*sum(r)/sum(r_max)
  return(rs)
}

# Choose the best cluster number
nC = 3:25 

cv.accuracy<- c()
for (i in 1:length(nC)) {
 
  coef = multinomialEM(training_data, nC[i])
  EM_predict = predictCF(validation_data, coef)
  cv.accuracy[i] <- R(EM_predict,validation_data,d=0,alpha=10)
  
}

#validation
# [1] 6.873618 7.745415 7.806073 7.449895 8.105736 7.695716 8.258315 7.049097 8.064468 8.284495
# [11] 7.937930 8.432793 7.996215 7.772650 8.306490 7.983393 8.634706 8.535889 7.486141 7.645274
# [21] 8.027794 8.946336 8.291748

plot(nC,cv.accuracy,type="b")


```

### Use the best parameter to calculate the accuracy
```{r}
# Test Error
best.C = 19
best.coef <- multinomialEM(train1, C = best.C)
best.pred <- predictCF(test1, best.coef)
accuracy1 <-sum(best.pred)/sum(test1)
accuracy2<-R(best.pred,test1,d=0,alpha=10)
# 39.4138

```


# Memory-based Algorithm


### similarity weights

##### Calculate pearson correlation
```{r}
weights_pearson <- cor(t(train1), method = "pearson")
#saveRDS(weights_pearson, file = "weights_pearson_train1.RData")


weights_pearson_train2 <- cor(t(train2), use = "pairwise.complete.obs", method = "pearson")
#saveRDS(weights_pearson_train2, file = "weights_pearson_train2.RData")

```

##### Calculate spearman correlation
```{r}
weights_spearman <- cor(t(train1), method = "spearman")
#saveRDS(weights_spearman, file = "weights_spearman_train1.RData")

weights_spearman_train2 <- cor(t(train2), use = "pairwise.complete.obs", method = "spearman")
#saveRDS(weights_spearman_train2, file = "weights_spearman_train2.RData")
```

##### Calculate Mean square difference
```{r}
meanSquareDiff <- function(df){
  m <- dim(df)[1] #user number
  dissim <- matrix(rep(NA, m * m), m, m)
  dissim <- data.frame(dissim)
  users <- rownames(df)
  colnames(dissim) <- users
  rownames(dissim) <- users
  for (i in 1:m){
    for (j in 1:m){
      r_i <- df[i,]
      r_j <- df[j,]
      dissim[i, j] <- mean((r_i - r_j)^2, na.rm = T)
    }
  }
  maxDissim <- max(dissim)
  sim <- (maxDissim - dissim)/maxDissim
  return (sim)
  
}

weights_MSD_train1 <- meanSquareDiff(train1)
#saveRDS(weights_MSD_train1, file = "weights_MSD_train1.RData")

weights_MSD_train2 <- meanSquareDiff(train2)
#saveRDS(weights_MSD_train2, file = "weights_MSD_train2.RData")

```

#### Calculate simrank matrix 
```{r}
compute_user_cum_movie_sum = function(m1,m2,
                                      Movie_Matrix_ = Movie_Matrix) {
  
  return(Movie_Matrix_[m1,m2])
}

compute_user_sim = function(j, k, iter = i,
                            train_adj_ = train_adj,
                            Movie_Matrix_ = Movie_Matrix,
                            Users_ = Users,
                            Movies_ = Movies) {
  
  if (j == k) {
    return(1)
  } else {
    moviesj = train_adj_$Movie[which(train_adj_$User == Users_[j])]
    moviesk = train_adj_$Movie[which(train_adj_$User == Users_[k])]
    
    # Pruning
    if (length(moviesj) + length(moviesk) - 
        length(unique(c(moviesj,moviesk))) <= 0.25 * length(moviesj)) {
      return(0)
    }
    
    # First Iter
    if (iter == 1) {
      return((length(moviesj) + length(moviesk) - 
                length(unique(c(moviesj,moviesk)))) * 0.8 / length(moviesk) / length(moviesj))
    }
    
    m1s = match(sort(rep(moviesj,length(moviesk))),Movies_)
    m2s = match(rep(moviesk,length(moviesj)),Movies_)
    
    mins = (m1s+m2s - sqrt((m1s-m2s)^2) )/2
    maxs = (m1s+m2s + sqrt((m1s-m2s)^2) )/2
    
    cum_sum = sum(mapply(compute_user_cum_movie_sum,
                         m1 = maxs, 
                         m2 = mins), na.rm=T)
    
    cum_sum = cum_sum * 0.8 / length(mins)
    return(cum_sum)
  }
}


compute_movie_cum_user_sum = function(m1,m2,
                                      User_Matrix_ = User_Matrix) {
  
  return(User_Matrix_[m1,m2])
}


compute_movie_sim = function(j, k, 
                             train_adj_ = train_adj,
                             User_Matrix_ = User_Matrix,
                             Users_ = Users,
                             Movies_ = Movies) {
  
  if (j == k) {
    return(1)
  } else {
    usersj = train_adj_$User[which(train_adj_$Movie == Movies_[j])]
    usersk = train_adj_$User[which(train_adj_$Movie == Movies_[k])]
    
    if (length(usersj) + length(usersk) - 
        length(unique(c(usersj,usersk))) <= 0.25 * length(usersj)) {
      return(0)
    }
    
    m1s = match(sort(rep(usersj,length(usersk))),Users_)
    m2s = match(rep(usersk,length(usersj)),Users_)
    
    mins = (m1s+m2s - sqrt((m1s-m2s)^2) )/2
    maxs = (m1s+m2s + sqrt((m1s-m2s)^2) )/2
    
    cum_sum = sum(mapply(compute_movie_cum_user_sum,
                         m1 = maxs, 
                         m2 = mins),na.rm=T)
    
    cum_sum = cum_sum * 0.8 / length(mins)
    return(cum_sum)
  }
}


generate_simrank_rdata = function(filename="simrank_matrix") {
  
  train = read.csv("../data/eachmovie_sample/data_train.csv")
  test = read.csv("../data/eachmovie_sample/data_test.csv")
  
  train$rescore = 0
  train$rescore[which(train$Score >= 6)] = 1
  
  train_adj = train[which(train$rescore ==1),]
  
  Movies = sort(unique(train_adj$Movie))
  Users = unique(train_adj$User)
  
  Movie_Matrix = diag(length(Movies))
  User_Matrix = diag(length(Users))
  
  for( i in 1:5 ) {
    print(paste0("i = ",i))
    for (j in 1:length(Users)) {
      if (j %% 100 == 0) {
        print(Sys.time())
        print(paste0("Users j=",j))
      }
      User_Matrix[j,1:j] = mapply(compute_user_sim,j=j,k=1:j)
    }
    
    for (j in 1:length(Movies)) {
      if (j %% 50 == 0) {
        print(Sys.time())
        print(paste0("Movies j=",j))
      }
      Movie_Matrix[j,1:j] = mapply(compute_movie_sim,j=j,k=1:j)
    }
    
  }
  
  User_Matrix[upper.tri(User_Matrix)] <- t(User_Matrix)[upper.tri(User_Matrix)]
  
 # save(User_Matrix,Users,file=paste0("../output/",filename,".RData"))
  
}
```

# Significance Weighting 

### Dataset 1 

```{r}
significance_weight_fn = function(a,u,df) {
  
  # i is the id of movie, a and u are id of users
  r_a = df[a, ]
  r_u = df[u, ] 
  
  r_a_mean = mean(r_a)
  r_a_sd = sd(r_a)
  r_u_mean = mean(r_u)
  r_u_sd = sd(r_u)
  
  num_a = which(r_a != 0)
  num_u = which(r_a != 0)
  
  n = length(intersect(num_a, num_u))
  significant_factor = ifelse(n < 50, n/50, 1)
  
  w_au = significant_factor * sum((r_a - r_a_mean) * (r_u - r_u_mean)) / (r_a_sd * r_u_sd)
  return(w_au)
  
}


users <- rownames(train1)
m <- length(users)

significant_weight <- matrix(rep(NA, m ^ 2), m, m)
significant_weight <- data.frame(significant_weight)

for (i in 1:m){
  for (j in 1:m){
    significant_weight[i,j] <- significance_weight_fn(users[i], users[j], train1)
  }
}

#saveRDS(significant_weight, file = "significant_weight Dataset1.RData")
```



### Dataset 2 

```{r}
significance_weight_fn2 = function(a,u,df) {
  a <- df[a, ]
  u <- df[u, ]
  
  NA_Col_a <- which(!is.na(a))
 
  NA_Col_u <- which(!is.na(u))
  
  c <- intersect(NA_Col_a,NA_Col_u)
  r_a <- as.numeric(a[ , c])
  r_u <- as.numeric(u[ , c])
  
  # i is the id of movie, a and u are id of users
  
  r_a_mean = mean(r_a)
  r_a_sd = sd(r_a)
  r_u_mean = mean(r_u)
  r_u_sd = sd(r_u)
  
  
  n <- length(c)
  significant_factor <- ifelse(n < 50, n/50, 1)
  
  w_au = significant_factor * sum((r_a - r_a_mean) %*% (r_u - r_u_mean)) / (r_a_sd * r_u_sd)
  return(w_au)
  
}


users <- rownames(train2)
m <- length(users)

significant_weight2 <- matrix(rep(NA, m * m), m, m)
significant_weight2 <- data.frame(significant_weight2)

for (i in 1:m){
  for (j in 1:m){
    significant_weight2[i,j] <- significance_weight_fn2(users[i], users[j], train2)
  }
}

#saveRDS(significant_weight2, file = "significant_weight Dataset2.RData")

```

# Variance weighting

### Dataset 1

```{r}

variance_weight_Fun = function(df) {
  
  # i is the id of movie, a and u are id of users

  
  n <- ncol(df)
  m <- nrow(df)
  var <- vector()
  
  for(i in 1:n) {
    var[i] <- sum((df[, i]-mean(df[,i]))^2)/(m - 1)
  }
  var_min <- min(var)
  var_max <- max(var)
  
  v_i <- vector()
  for(i in 1:n) {
    v_i[i] <- (var[i] - var_min) / var_max
  }
  
  
  variance_weight <- matrix(rep(NA, m * m), m, m)
  variance_weight <- data.frame(variance_weight)
  
  for (a in 1:m){
   for (u in 1:m){
    r_a = df[a, ]
    r_u = df[u, ] 
  
    r_a_mean = mean(r_a)
    r_a_sd = sd(r_a)
    r_u_mean = mean(r_u)
    r_u_sd = sd(r_u)
  
    variance_weight[a,u] <- sum(((r_a - r_a_mean)/r_a_sd) * ((r_u - r_u_mean)/r_u_sd) * v_i) / sum(v_i)
  }
}
  
  return(variance_weight)
  
}

variance_weight1<-variance_weight_Fun(train1)
#saveRDS(variance_weight1, file = "variance_weight Dataset1.RData")
```


### Dataset 2

```{r}

variance_weight_fun2= function(a,u,df) {
  
  a <- df[a, ]
  u <- df[u, ]
  
  NA_Col_a <-  which(!is.na(a))
  NA_Col_u <-  which(!is.na(a))
  
  c <- intersect(NA_Col_a,NA_Col_u)
  r_a <- as.numeric(a[ , c])
  r_u <- as.numeric(u[ , c])
  # i is the id of movie, a and u are id of users
  
  r_a_mean = mean(r_a)
  r_a_sd = sd(r_a)
  r_u_mean = mean(r_u)
  r_u_sd = sd(r_u)
  
  
  n = length(c)
  var <- vector()
  for(i in 1:n) {
    var[i] <- sum((na.omit(df[,c(i)])-mean(na.omit(df[,c(i)])))^2)/(length(na.omit(df[,c(i)])) - 1)
  }
  var_min <- min(var)
  var_max <- max(var)
  
  v_i <- vector()
  for(i in 1:n) {
    v_i[i] <- (sum((na.omit(df[,c(i)])-mean(na.omit(df[,c(i)])))^2/(length(na.omit(df[,c(i)])) - 1)) - var_min) / var_max
  }
  
  w_au = sum((r_a - r_a_mean) %*% (r_u - r_u_mean) %*% v_i) / sum(v_i)
  return(w_au)
  
}


users <- rownames(train2)
m <- length(users)

variance_weight2 <- matrix(rep(NA, m * m), m, m)
variance_weight2 <- data.frame(variance_weight2)

for (i in 1:m){
  for (j in 1:m){
    user_i <- users[i]
    user_j <- users[j]
    tmp  = train2[user_i, user_j]
    if (is.integer(tmp)){
      variance_weight2[i,j] <- variance_weight_fun2(users[i], users[j], train2)
    }
  }
}

#saveRDS(variance_weight2, file = "variance_weight Dataset2.RData") 

```

# Select neighbors: best-n

```{r}

selectNeighborWeights <- function(df, weights, n){
  users <- rownames(df)
  m <- length(users)
  neighbors <- matrix(rep(NA, m * n), m, n)
  neighors <- data.frame(neighbors)
  rownames(neighbors) <- users
  
  for (i in 1:m){
    neighbors[i, ] <- colnames(weights)[order(weights[i, ], decreasing=TRUE)][1:n]
  }
  return (neighbors)
}

```


# Rating Normalization - z-score
```{r}

computeZScore <- function(weights, neighbors, df, a, i){
  neighbors_a <- neighbors[toString(a), ] 
  num <- rep(0, length(neighbors_a))
  for (u in 1:length(neighbors_a)){
    neighborU <- neighbors_a[u] 
    r_ui <- df[toString(neighborU), toString(i)]
    r_u <- mean(df[neighborU, ], na.rm = T)
    sd_u <- sd(df[neighborU, ], na.rm = F)
    w_au <- weights[toString(a), neighborU]
    num[u] <- (r_ui - r_u)/sd_u * w_au
  }
  w_a <- sum(weights[toString(a), neighbors_a])
  sd_a <- sd(df[toString(a), ])
  index_not_i <- colnames(df) != i
  r_a <- mean(df[toString(a), index_not_i])
  p_ai <- r_a + sd_a * sum(num)/w_a
  return (p_ai)
}

computeZScoreMatrix <- function(weights, neighbors, df, df_test){
  a <- dim(df_test)[1]
  b <- dim(df_test)[2]
  users <- rownames(df_test)
  website <- colnames(df_test)
  computeZScoreMatrix <- matrix(rep(0, a * b), a, b)
  for (i in 1:a){
    for (j in 1:b){
      ##if (is.na(df_test[i, j]) == F){ #data2
        if (df_test[i, j] != 0){ data1
        computeZScoreMatrix[i, j] <- computeZScore(weights, neighbors, df, users[i], website[j])
      }
      #computeZScoreMatrix[i, j] <- computeZScore(weights, neighbors, df, users[i], website[j])
    }
  }
  return (computeZScoreMatrix)
}

```

## Predict


#####  Pearson

```{r}
neighbors_pearson_train1 <- selectNeighborWeights(train1, weights_pearson, 20)
neighbors_pearson_train2 <- selectNeighborWeights(train2, weights_pearson_train2, 20)
ZScoreMatrix_pearson_test1 <- computeZScoreMatrix(weights_pearson, neighbors_pearson_train1, train1,test1)
ZScoreMatrix_pearson_test2 <- computeZScoreMatrix(weights_pearson_train2, neighbors_pearson_train2, train2,test2)
```

##### Spearman

```{r}
neighbors_spearman_train1 <- selectNeighborWeights(train1, weights_spearman, 20)
neighbors_spearman_train2 <- selectNeighborWeights(train2, weights_spearman_train2, 20)
ZScoreMatrix_spearman_test1 <- computeZScoreMatrix(weights_spearman, neighbors_spearman_train1, train1,test1)
ZScoreMatrix_spearman_test2 <- computeZScoreMatrix(weights_spearman_train2, neighbors_spearman_train2, train2,test2)
```

##### MSD

```{r}
neighbors_MSD_train1 <- selectNeighborWeights(train1, weights_MSD_train1, 20)
neighbors_MSD_train2 <- selectNeighborWeights(train2, weights_MSD_train2, 20)
ZScoreMatrix_MSD_test1 <- computeZScoreMatrix(weights_MSD_train1, neighbors_MSD_train1, train1,test1)
ZScoreMatrix_MSD_test2 <- computeZScoreMatrix(weights_MSD_train2, neighbors_MSD_train2, train2,test2)
```


##### SimRank
```{r}
neighbors_SR_train1 <- selectNeighborWeights(train1, User_Matrix, 20)
neighbors_SR_train2 <- selectNeighborWeights(train2, Movie_Matrix, 20)
ZScoreMatrix_SR_test1 <- computeZScoreMatrix(User_Matrix, neighbors_SR_train1, train1,test1)
ZScoreMatrix_SR_test2 <- computeZScoreMatrix(Movie_Matrix, neighbors_SR_train2, train2,test2)
```

##### Pearson + Sig

```{r}
neighbors_pearson_sig_train1 <- selectNeighborWeights(train1, significant_weight, 20)
neighbors_pearson_sig_train2 <- selectNeighborWeights(train2, significant_weight2, 20)
ZScoreMatrix_pearson_sig_test1 <- computeZScoreMatrix(significant_weight, neighbors_pearson_sig_train1, train1,test1)
ZScoreMatrix_pearson_sig_test2 <- computeZScoreMatrix(significant_weight2, neighbors_pearson_sig_train2, train2,test2)
```


##### Pearson + Var

```{r}
neighbors_pearson_var_train1 <- selectNeighborWeights(train1, variance_weight1, 20)
neighbors_pearson_var_train2 <- selectNeighborWeights(train2, variance_weight2, 20)
ZScoreMatrix_pearson_var_test1 <- computeZScoreMatrix(variance_weight1, neighbors_pearson_var_train1, train1,test1)
ZScoreMatrix_pearson_var_test2 <- computeZScoreMatrix(variance_weight2, neighbors_pearson_var_train2, train2,test2)
```




